Correcting for bias in high throughput sequencing data

### 2011-06-12 seqbias-paper.tex
compiling in bioinformatics template

\documentclass[letterpaper]{article}
\usepackage{qcjstd}
\usepackage{graphicx}
\usepackage{lscape}
\usepackage{multicol}
\usepackage{geometry}
\usepackage{amsmath}
\usepackage{multirow}
%\usepackage{bookman}
%\usepackage{cmbright}

\makeatletter
\newenvironment{tablehere}
  {\def\@captype{table}}
  {\vspace{1em}}

\newenvironment{figurehere}
  {\def\@captype{figure}}
  {\vspace{1em}}
\makeatother

\geometry{left=0.8in,right=0.8in, top=1in, bottom=1in}

\title{A Model for Sequence Bias in RNA-Seq}

\begin{document}

%\maketitle

\qcjtitle{\sc{(Draft)}}{\sc{A Model for Sequencing Bias in RNA-Seq}}{Daniel C. Jones}

@@

\begin{multicols}{2}

@@

\begin{figure*}
\begin{center}
\includegraphics[width=0.75\textwidth]{fig/freqs.pdf}
\end{center}
\caption{Nucleotide frequencies relative to the start of each read (labeled
position 0), respecting strand. The divergence of the nucleotide distribution
is summarized using the symmetrized Kullback-Leibler divergence (see Section
3).}
\label{fig:freqs}
\end{figure*}

@@

\begin{figure*}
\begin{center}
\includegraphics[width=0.75\textwidth]{fig/structure2.pdf}
\end{center}
\caption{The network structure learned on the Trapnell data set, where 0 is the
read start. Dotted boxes represent positions not included in the model (i.e.
positions deemed uninformative).}
\label{fig:structure}
\end{figure*}

@@

\begin{tablehere}
\begin{center}
\footnotesize{
\begin{tabular}{lllr}
\textbf{Experiment} & \textbf{Species} & \textbf{Platform} & \textbf{Read Length} \\ \hline
Wetterbom \cite{Wetterbom2010} & Chimpanzee & ABI & 33 \\
Katze (unpublished) & Macaque & ABI & 50 \\
Bullard \cite{Bullard2010} & Human & Illumina & 35 \\
Mortazavi \cite{Mortazavi2008} & Mouse & Illumina & 33 \\
Trapnell \cite{Trapnell2010} & Mouse & Illumina & 75
\end{tabular}
}
\end{center}
\caption{Data sets on which the methods are evaluated.}
\label{tab:datasets}
\end{tablehere}

@@

\begin{figurehere}
\begin{center}
\includegraphics[width=0.5\textwidth]{fig/kl.pdf}
\end{center}
\caption{Plots of the KL divergence over overlapping $k$-mers in the Trapnell
experiment.}
\label{fig:kl}
\end{figurehere}

@@

\begin{figurehere}
\begin{center}
\includegraphics[width=0.35\textwidth]{fig/kl-delta.pdf}
\end{center}
\caption{Summed KL divergence, normalized by unadjusted divergence, so that a
number less than 1 indicates a decrease in divergence, or an increase in
uniformity.}
\label{fig:kl-delta}
\end{figurehere}

@@

\begin{tablehere}
\begin{center}
\tiny{
\begin{tabular}{lrrrrrr}
         &\textbf{Avg-7mer}&\textbf{7mer}&\textbf{4mer}&\textbf{GLM}&\textbf{MART}&\textbf{BN} \\
\textbf{Katze}            &  0.004  & 0.017  & 0.055 &0.166  &0.248&0.305 \\
\textbf{Wetterbom}        & -0.076  &-0.058  &-0.002 &0.240  &0.070&0.557 \\
\textbf{Bullard}          &  0.124  & 0.192  & 0.139 &0.285  &0.198&0.380 \\
\textbf{Mortazavi}        &  0.059  & 0.123  & 0.099 &0.284  &0.310&0.401 \\
\textbf{Trapnell}         &  0.094  & 0.163  & 0.104 &0.265  &0.332&0.349 \\
\end{tabular}
}
\end{center}
\caption{Change in fit, compared to the null model of uniform Poisson rate across exons.}
\label{tab:pois}
\end{tablehere}

@@

\begin{figurehere}
\begin{center}
\includegraphics[width=0.4\textwidth]{fig/pois.pdf}
\end{center}
\caption{Plotted is the change in fit of each of the test exons. A
positive value indicates a better fit in the bias corrected model.}
\label{fig:pois}
\end{figurehere}

@@

\end{multicols}

>>>>

\documentclass{bioinfo}
\copyrightyear{2011}
\pubyear{2011}

\usepackage{comment}

\begin{document}
\firstpage{1}

\title{Correcting for bias in high throughput sequencing data}
\author[Jones \textit{et~al}]
{Daniel C. Jones\,$^{1,}$
\footnote{to whom correspondence should be addressed}\hspace{0.5em}
Walter L. Ruzzo\,$^{1,2}$
Xinxia Peng\,$^{3}$
Michael G. Katze$^{3}$
}

\history{Received on XXXXX; revised on XXXXX; accepted on XXXXX}

\address{
$^{1}$Deportment of Computer Science and Engineering, University of
Washington, Seattle, WA 98195-2350, USA\\
$^{2}$Fred Hutchinson Cancer Research Center, Seattle, WA 98109\\
$^{3}$Department of Microbiology, University of Washington, Seattle, WA
98195-7242, USA}

\editor{Associate Editor: XXXXXXX}

\maketitle

@@

@@

\begin{comment}
\begin{figure*}
\begin{center}
\includegraphics[width=0.75\textwidth]{fig/freqs.pdf}
\end{center}
\caption{Nucleotide frequencies relative to the start of each read (labeled
position 0), respecting strand. The divergence of the nucleotide distribution
is summarized using the symmetrized Kullback-Leibler divergence (see Section
3).}
\label{fig:freqs}
\end{figure*}
\end{comment}

@@

\begin{comment}
\begin{figure*}
\begin{center}
\includegraphics[width=0.75\textwidth]{fig/structure2.pdf}
\end{center}
\caption{The network structure learned on the Trapnell data set, where 0 is the
read start. Dotted boxes represent positions not included in the model (i.e.
positions deemed uninformative).}
\label{fig:structure}
\end{figure*}
\end{comment}

@@

%% TODO
\begin{comment}
\begin{tablehere}
\begin{center}
\footnotesize{
\begin{tabular}{lllr}
\textbf{Experiment} & \textbf{Species} & \textbf{Platform} & \textbf{Read Length} \\ \hline
Wetterbom \cite{Wetterbom2010} & Chimpanzee & ABI & 33 \\
Katze (unpublished) & Macaque & ABI & 50 \\
Bullard \cite{Bullard2010} & Human & Illumina & 35 \\
Mortazavi \cite{Mortazavi2008} & Mouse & Illumina & 33 \\
Trapnell \cite{Trapnell2010} & Mouse & Illumina & 75
\end{tabular}
}
\end{center}
\caption{Data sets on which the methods are evaluated.}
\label{tab:datasets}
\end{tablehere}
\end{comment}

@@

%% TODO
\begin{comment}
\begin{figurehere}
\begin{center}
\includegraphics[width=0.5\textwidth]{fig/kl.pdf}
\end{center}
\caption{Plots of the KL divergence over overlapping $k$-mers in the Trapnell
experiment.}
\label{fig:kl}
\end{figurehere}
\end{comment}

@@

%% TODO
\begin{comment}
\begin{figurehere}
\begin{center}
\includegraphics[width=0.35\textwidth]{fig/kl-delta.pdf}
\end{center}
\caption{Summed KL divergence, normalized by unadjusted divergence, so that a
number less than 1 indicates a decrease in divergence, or an increase in
uniformity.}
\label{fig:kl-delta}
\end{figurehere}
\end{comment}

@@

\begin{comment}
\begin{tablehere}
\begin{center}
\tiny{
\begin{tabular}{lrrrrrr}
         &\textbf{Avg-7mer}&\textbf{7mer}&\textbf{4mer}&\textbf{GLM}&\textbf{MART}&\textbf{BN} \\
\textbf{Katze}            &  0.004  & 0.017  & 0.055 &0.166  &0.248&0.305 \\
\textbf{Wetterbom}        & -0.076  &-0.058  &-0.002 &0.240  &0.070&0.557 \\
\textbf{Bullard}          &  0.124  & 0.192  & 0.139 &0.285  &0.198&0.380 \\
\textbf{Mortazavi}        &  0.059  & 0.123  & 0.099 &0.284  &0.310&0.401 \\
\textbf{Trapnell}         &  0.094  & 0.163  & 0.104 &0.265  &0.332&0.349 \\
\end{tabular}
}
\end{center}
\caption{Change in fit, compared to the null model of uniform Poisson rate across exons.}
\label{tab:pois}
\end{tablehere}
\end{comment}

@@

%% TODO
\begin{comment}
\begin{figurehere}
\begin{center}
\includegraphics[width=0.4\textwidth]{fig/pois.pdf}
\end{center}
\caption{Plotted is the change in fit of each of the test exons. A
positive value indicates a better fit in the bias corrected model.}
\label{fig:pois}
\end{figurehere}
\end{comment}

@@
